|

### 2024-09-04 manuscript/paperCMSACversion.Rmd
updcated.

summ <-
  dat %>% group_by(EF) %>% summarize(
    n = n(),
    mnage = round(mean(age), 2),
    seage = round(sd(age) / sqrt(n),3),
    mnwt = round(mean(Wt), 2),
    sewt = round(sd(Wt) / sqrt(n),3),
    mnht = round(mean(Ht), 2),
    seht = round(sd(Ht) / sqrt(n),3),
    mngoals = round(mean(G), 2),
    segoals = round(sd(G) / sqrt(n),3),
    mntf = round(mean(Total.Fights), 2),
    setf = round(sd(Total.Fights) / sqrt(n),3),
    mnpimgp = round(mean(PIM.GP_numeric), 2),
    sepimgp = round(sd(PIM.GP_numeric) / sqrt(n),3)

@@

summEP <-
  dat %>% group_by(EP) %>% summarize(
    n = n(),
    mnage = round(mean(age), 2),
    seage = round(sd(age) / sqrt(n),3),
    mnwt = round(mean(Wt), 2),
    sewt = round(sd(Wt) / sqrt(n),3),
    mnht = round(mean(Ht), 2),
    seht = round(sd(Ht) / sqrt(n),3),
    mngoals = round(mean(G), 2),
    segoals = round(sd(G) / sqrt(n),3),
    mntf = round(mean(Total.Fights), 2),
    setf = round(sd(Total.Fights) / sqrt(n),3),
    mnpimgp = round(mean(PIM.GP_numeric), 2),
    sepimgp = round(sd(PIM.GP_numeric) / sqrt(n),3)
  )

@@

However, we believe that some of the methods used in @Popkin2023 to reach these conclusions are not statistically appropriates.  Most notably, the method used to reached their most sensational finding - enforcers died on average about 10 years earlier than their non-enforcer counterparts - is statistically unsuitable. Specifically, to reach this conclusion, they use the data from the matched cohort study but then condition on death.  This effectively breaks the matching that they appropriately used to compare mortality rates, and therefore this conclusion is not statistically valid.

Here, we attempt not to replicate the results of the analysis in @Popkin2023, but what we believe to be more appropriate statistical analysis, specifically survival analysis methodology. This shift in perspective allows us to overcome the weakness of mortality rate, which is that it is a function of when it is calculated (i.e. In all cases, mortality rate will be 100% when all members of the original sample have died). Instead, survival analysis provides a survival curve, which can be used to analyze how the probability of survival changes over time.  In addition, we augmented the data used in @Popkin2023 to increase the sample size in the analysis and include more deceased players in the data set.

@@

<!-- <!-- (Greg comment: After matching you CAN compare mortality rates and that comparison would be valid AT THAT GIVEN POINT when they are compared.  But a better analysis would be to use survival analysis.  The problem with mortality rate is that eventually everyone in both groups will be dead if you wait long enough and that comparison is meaningless. Survival times are much better comparison. ) --> -->

@@

# Data 
## Data from orignal paper
We obtained the two data sets used in @Popkin2023: general table (n = 7,432) and death table (n = 45).  As in @Popkin2023, we considered only players who's first season was 1967 or later and played any games between the 1967-68 and the 2021-22 NHL seasons.  This leaves 6,039 players after filtering (the exact number mentioned in @Popkin2023).  We then removed 4 players with missing height, leaving us with `r nrow(dat)` (6,035) remaining observations, of which 252 are deaths, for our re-analysis.

@@

We note than in @Popkin2023, they find 183 E-P enforcers; we have only 182 in our data set (check that this is because of the missing height issue.)

@@

Of the players included in the analysis, on average, enforcers in this data set tended to be older than the non-enforcers (EF: `r summ$mnage[summ$EF]` vs non-EF: `r summ$mnage[!summ$EF]` and EP: `r summEP$mnage[summEP$EP]` vs non-EP: `r summEP$mnage[!summEP$EP]`).  Additionally, as seen in Tables \ref{tab:EDAEF} and \ref{tab:EDAEP} enforcers tended to also be taller and heavier on average than non-enforcers. The means of age, weight, height and goals are all significantly different between the enforcers and non-enforcer groups for both definitions of enforcers.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|}
\hline
 Variable & mean (sd)\\
 \hline
 n & `r nrow(dat)`    \\
 n deceased & `r sum(dat$dead)` (`r paste0(round(100*sum(dat$dead)/nrow(dat),2),"\\%")`)    \\
 Age (Years) & `r round(mean(as.numeric(dat$age)),2)` (`r round(sd(as.numeric(dat$age)),2)`)\\
 Height (Inches) & `r round(mean(as.numeric(dat$Ht)),2)` (`r round(sd(as.numeric(dat$Ht)),2)`)\\
 Weight (Pounds) & `r mean(as.numeric(dat$Wt))` (`r sd(as.numeric(dat$Wt))`)\\
 Fights per game & `r mean(as.numeric(dat$FperG))` (`r sd(as.numeric(dat$FperG))`)\\
 Penalty Minutes per game  & `r mean(as.numeric(dat$PIM.GP_numeric))`  (`r  sd(as.numeric(dat$PIM.GP_numeric))`)\\
 \hline
\end{tabular}
\end{center}
\caption{Overall summary stats \label{tab:EDAoverall}}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 Variable & EF = TRUE & EF = FALSE & p-value \\
 \hline
 n & `r summ$n[summ$EF]` & `r summ$n[!summ$EF]` & \\
 Age (years) & `r summ$mnage[summ$EF]` (`r summ$seage[summ$EF]`) & `r summ$mnage[!summ$EF]` (`r summ$seage[!summ$EF]`) & `r ifelse(round(t.test(dat$age~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EF)$p.value,2))`\\
 Weight (lbs) & `r summ$mnwt[summ$EF]` (`r summ$sewt[summ$EF]`) & `r summ$mnwt[!summ$EF]` (`r summ$sewt[!summ$EF]`) & `r ifelse(round(t.test(dat$Wt~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EF)$p.value,2))`\\
  Height (in) & `r summ$mnht[summ$EF]` (`r summ$seht[summ$EF]`) & `r summ$mnht[!summ$EF]` (`r summ$seht[!summ$EF]`) & `r ifelse(round(t.test(dat$Ht~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$Ht~dat$EF)$p.value,2))`\\
    Goals  & `r summ$mngoals[summ$EF]` (`r summ$segoals[summ$EF]`) & `r summ$mngoals[!summ$EF]` (`r summ$segoals[!summ$EF]`) & `r ifelse(round(t.test(dat$G~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EF)$p.value,2))`\\
 \hline
\end{tabular}
\end{center}
\caption{Comparing enforcers to non enforcers when enforcer is defined as E-F \label{tab:EDAEF}}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 Variable & EP = TRUE & EP = FALSE & p-value \\
 \hline
 n & `r summEP$n[summEP$EP]` & `r summEP$n[!summEP$EP]` & \\
  GP & `r summEP$mnage[summEP$EP]` (`r summEP$seage[summEP$EP]`) & `r summEP$mnage[!summEP$EP]` (`r summEP$seage[!summEP$EP]`) & `r ifelse(round(t.test(dat$age~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EP)$p.value,2))`\\
 Age (years) & `r summEP$mnage[summEP$EP]` (`r summEP$seage[summEP$EP]`) & `r summEP$mnage[!summEP$EP]` (`r summEP$seage[!summEP$EP]`) & `r ifelse(round(t.test(dat$age~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EP)$p.value,2))`\\
 Weight (lbs) & `r summEP$mnwt[summEP$EP]` (`r summEP$sewt[summEP$EP]`) & `r summEP$mnwt[!summEP$EP]` (`r summEP$sewt[!summEP$EP]`) & `r ifelse(round(t.test(dat$Wt~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EP)$p.value,2))`\\
  Height (in) & `r summEP$mnht[summEP$EP]` (`r summEP$seht[summEP$EP]`) & `r summEP$mnht[!summEP$EP]` (`r summEP$seht[!summEP$EP]`) & `r ifelse(round(t.test(dat$Ht~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$Ht~dat$EP)$p.value,2))`\\
      Goals  & `r summEP$mngoals[summEP$EP]` (`r summEP$segoals[summEP$EP]`) & `r summEP$mngoals[!summEP$EP]` (`r summEP$segoals[!summEP$EP]`) & `r ifelse(round(t.test(dat$G~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EP)$p.value,2))`\\
 \hline
\end{tabular}
\end{center}

@@

# Methods {#sec:methods}
The analysis presented here utilizes all hockey players in the data, both alive and deceased, in a survival analysis to explore differences in survival times between enforcers an non-enforcers, which we believe to be a more appropriate method for analyizing the problem in this setting.  Rather than focusing on mortality rates, as in @popkin2023, we instead focus on survival times and leverage survival analysis techniques, specifically Cox proportional hazards models, to look for differences in survival times of enforcers versus non-enforcers.

@@

>>>>

dat$Pos2 <- as.character(dat$Pos)
dat$Pos2[dat$Pos2 %in% c("L","R")] <- "W"
dat$Pos2 <- factor(dat$Pos2)

summ <-
  dat %>% group_by(EF) %>% summarize(
    n = n(),
    mnGP = round(mean(GP.x), 2),
    seGP = round(sd(GP.x), 2),
    mnage = round(mean(age), 2),
    seage = round(sd(age) / sqrt(n),3),
    mnwt = round(mean(Wt), 2),
    sewt = round(sd(Wt) / sqrt(n),3),
    mnht = round(mean(Ht), 2),
    seht = round(sd(Ht) / sqrt(n),3),
    mngoals = round(mean(G), 2),
    segoals = round(sd(G) / sqrt(n),3),
    mntf = round(mean(Total.Fights), 2),
    setf = round(sd(Total.Fights) / sqrt(n),3),
    mnpimgp = round(mean(PIM.GP_numeric), 2),
    sepimgp = round(sd(PIM.GP_numeric) / sqrt(n),3),
    pD = round(mean(Pos2 == "D"),3),
    pW = round(mean(Pos2 == "W"),3),
    pC = round(mean(Pos2 == "C"),3)

@@

summEP <-
  dat %>% group_by(EP) %>% summarize(
    n = n(),
    mnGP = round(mean(GP.x), 2),
    seGP = round(sd(GP.x), 2),
    mnage = round(mean(age), 2),
    seage = round(sd(age) / sqrt(n),3),
    mnwt = round(mean(Wt), 2),
    sewt = round(sd(Wt) / sqrt(n),3),
    mnht = round(mean(Ht), 2),
    seht = round(sd(Ht) / sqrt(n),3),
    mngoals = round(mean(G), 2),
    segoals = round(sd(G) / sqrt(n),3),
    mntf = round(mean(Total.Fights), 2),
    setf = round(sd(Total.Fights) / sqrt(n),3),
    mnpimgp = round(mean(PIM.GP_numeric), 2),
    sepimgp = round(sd(PIM.GP_numeric) / sqrt(n),3),
    pD = round(mean(Pos2 == "D"),3),
    pW = round(mean(Pos2 == "W"),3),
    pC = round(mean(Pos2 == "C"),3)
  )

@@

<!-- Most notably, the method used to reached their most sensational finding - enforcers died on average about 10 years earlier than their non-enforcer counterparts - is statistically unsuitable. Specifically, to reach this conclusion, they use the data from the matched cohort study but then condition on death.  This effectively breaks the matching that they appropriately used to compare mortality rates, and therefore this conclusion is not statistically valid.   -->

However, we believe that some of the statistical methods used in @Popkin2023 to reach these conclusions are not the best choices given the problem that is studied.  Here, we attempt not to replicate the results of the analysis in @Popkin2023, but what we believe to be more appropriate statistical methods, specifically survival analysis techniques. This shift in perspective allows us to overcome the weakness of mortality rate, which is that it is a function of when it is calculated (i.e. In all cases, mortality rate will be 100% when all members of the original sample have died). Instead, survival analysis provides a survival curve, which can be used to analyze how the probability of survival changes over time.  In addition, we augmented the data used in @Popkin2023 to increase the sample size in the analysis and include more deceased players in the data set.

@@

<!-- <!-- (Greg comment: After matching you CAN compare mortality rates and that comparison would be valid AT THAT GIVEN POINT when they are compared.  But a better analysis would be to use survival analysis.  The problem with mortality rate is that eventually everyone in both groups will be dead if you wait long enough and that comparison is meaningless. Survival times are much better comparison. ) -->

@@

# Data 
## Data from original paper
We obtained the two data sets used in @Popkin2023: general table (n = 7,432) and death table (n = 45).  As in @Popkin2023, we considered only players who's first season was 1967 or later and played any games between the 1967-68 and the 2021-22 NHL seasons.  This leaves 6,039 players after filtering (the exact number mentioned in @Popkin2023).  We then removed 4 players with missing height, leaving us with `r nrow(dat)` (6,035) remaining observations, of which 252 are deaths, for our re-analysis.

@@

We note than in @Popkin2023, they find 183 E-P enforcers; we have only 182 in our data set with the discrepancy due the the removal of the 4 players with missing height information.

@@

<!-- This is what they matched on " including date of birth, total number of games played, height, weight, and position played." -->

In this same set of players, on average, enforcers in this data set tended to be older than the non-enforcers (EF: `r summ$mnage[summ$EF]` vs non-EF: `r summ$mnage[!summ$EF]` and EP: `r summEP$mnage[summEP$EP]` vs non-EP: `r summEP$mnage[!summEP$EP]`).  Additionally, as seen in Tables \ref{tab:EDAEF} and \ref{tab:EDAEP} enforcers tended to also be taller and heavier on average than non-enforcers. The means of age, weight, height and played are all significantly different between the enforcers and non-enforcer groups for both definitions of enforcers.  We also note differences some substantial differences in the positional distribution of enforcers vs non-enforcers.  Specifically, the percentage of defensive players in the enforcer vs non-enforcers is relatively close there are very few enforcers, defined with either E-F or E-P, who are listed as playing center.  This also means that the percentage of enforcers who are listed as wings (either right or left) is higher for enforcers than for non-enforcers.  The variables listed here are the same as those that were used for matching in @Popkin2023 (i.e. date of birth, total number of games played, height, weight and position).

<!-- Goals  & `r summ$mngoals[summ$EF]` (`r summ$segoals[summ$EF]`) & `r summ$mngoals[!summ$EF]` (`r summ$segoals[!summ$EF]`) & `r ifelse(round(t.test(dat$G~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EF)$p.value,2))`\\ -->

\begin{table}
\begin{center}
\begin{tabular}{|c|c|}
\hline
 Variable & mean (sd)\\
 \hline
 n & `r nrow(dat)`    \\
 n deceased & `r sum(dat$dead)` (`r paste0(round(100*sum(dat$dead)/nrow(dat),2),"\\%")`)    \\
 Games Played & `r  round(mean(dat$GP.x),2)` (`r round(sd(dat$GP.x),2)`)\\
 Age (Years) & `r round(mean(as.numeric(dat$age)),2)` (`r round(sd(as.numeric(dat$age)),2)`)\\
 Height (Inches) & `r round(mean(as.numeric(dat$Ht)),2)` (`r round(sd(as.numeric(dat$Ht)),2)`)\\
 Weight (Pounds) & `r mean(as.numeric(dat$Wt))` (`r sd(as.numeric(dat$Wt))`)\\
 Fights per game & `r mean(as.numeric(dat$FperG))` (`r sd(as.numeric(dat$FperG))`)\\
 Penalty Minutes per game  & `r mean(as.numeric(dat$PIM.GP_numeric))`  (`r  sd(as.numeric(dat$PIM.GP_numeric))`)\\
 \hline
\end{tabular}
\end{center}
\caption{Overall summary stats \label{tab:EDAoverall}}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 Variable & EF = TRUE & EF = FALSE & p-value \\
 \hline
 n & `r summ$n[summ$EF]` & `r summ$n[!summ$EF]` & \\
 Games Played & `r summ$mnGP[summ$EF]` (`r summ$seGP[summ$EF]`) & `r summ$mnGP[!summ$EF]` (`r summ$seGP[!summ$EF]`) & `r ifelse(round(t.test(dat$GP.x~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$GP.x~dat$EF)$p.value,2))`\\
 Age (years) & `r summ$mnage[summ$EF]` (`r summ$seage[summ$EF]`) & `r summ$mnage[!summ$EF]` (`r summ$seage[!summ$EF]`) & `r ifelse(round(t.test(dat$age~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EF)$p.value,2))`\\
 Weight (lbs) & `r summ$mnwt[summ$EF]` (`r summ$sewt[summ$EF]`) & `r summ$mnwt[!summ$EF]` (`r summ$sewt[!summ$EF]`) & `r ifelse(round(t.test(dat$Wt~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EF)$p.value,2))`\\
  Height (in) & `r summ$mnht[summ$EF]` (`r summ$seht[summ$EF]`) & `r summ$mnht[!summ$EF]` (`r summ$seht[!summ$EF]`) & `r ifelse(round(t.test(dat$Ht~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$Ht~dat$EF)$p.value,2))`\\
          percent D & `r paste0(100*summ$pD[summ$EF],"\\%")` & `r paste0(100*summ$pD[!summ$EF],"\\%")` & \\
    percent W & `r paste0(100*summ$pW[summ$EF],"\\%")` & `r paste0(100*summ$pW[!summ$EF],"\\%")` & \\
    percent C & `r paste0(100*summ$pC[summ$EF],"\\%")` & `r paste0(100*summ$pC[!summ$EF],"\\%")` & \\

 \hline
\end{tabular}
\end{center}
\caption{Comparing enforcers to non enforcers when enforcer is defined as E-F \label{tab:EDAEF}}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 Variable & EP = TRUE & EP = FALSE & p-value \\
 \hline
 n & `r summEP$n[summEP$EP]` & `r summEP$n[!summEP$EP]` & \\
 Games Played & `r summEP$mnGP[summEP$EP]` (`r summEP$seGP[summEP$EP]`) & `r summEP$mnGP[!summEP$EP]` (`r summEP$seGP[!summEP$EP]`) & `r ifelse(round(t.test(dat$GP.x~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$GP.x~dat$EP)$p.value,2))`\\
 Age (years) & `r summEP$mnage[summEP$EP]` (`r summEP$seage[summEP$EP]`) & `r summEP$mnage[!summEP$EP]` (`r summEP$seage[!summEP$EP]`) & `r ifelse(round(t.test(dat$age~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EP)$p.value,2))`\\
 Weight (lbs) & `r summEP$mnwt[summEP$EP]` (`r summEP$sewt[summEP$EP]`) & `r summEP$mnwt[!summEP$EP]` (`r summEP$sewt[!summEP$EP]`) & `r ifelse(round(t.test(dat$Wt~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$age~dat$EP)$p.value,2))`\\
  Height (in) & `r summEP$mnht[summEP$EP]` (`r summEP$seht[summEP$EP]`) & `r summEP$mnht[!summEP$EP]` (`r summEP$seht[!summEP$EP]`) & `r ifelse(round(t.test(dat$Ht~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$Ht~dat$EP)$p.value,2))`\\
    percent D & `r paste0(100*summEP$pD[summEP$EP],"\\%")` & `r paste0(100*summEP$pD[!summEP$EP],"\\%")` & \\
    percent W & `r paste0(100*summEP$pW[summEP$EP],"\\%")` & `r paste0(100*summEP$pW[!summEP$EP],"\\%")` & \\
    percent C & `r paste0(100*summEP$pC[summEP$EP],"\\%")` & `r paste0(100*summEP$pC[!summEP$EP],"\\%")` & \\
 \hline
\end{tabular}
\end{center}

@@

<!-- ```{r KMEP, echo = FALSE, fig.cap="Kaplan Meier plot for EP" } -->
<!-- dat <- dat %>% mutate(northamerica = ifelse(Ctry %in% c("USA","CAN"),"North America","Other")) -->
<!-- library(ggsurvfit) -->
<!-- survfit2(Surv ~ northamerica, data = dat) %>%  -->
<!--   ggsurvfit() + -->
<!--   labs( -->
<!--     x = "Age", -->
<!--     y = "Overall survival probability" -->
<!--   ) +  -->
<!--   add_confidence_interval() -->
<!-- ``` -->

# Methods {#sec:methods}
The analysis presented here utilizes all hockey players in the data, both alive and deceased, in a survival analysis to explore differences in survival times between enforcers an non-enforcers, which we believe to be a more appropriate method for analyizing the problem in this setting.  Rather than focusing on mortality rates, as in @Popkin2023, we instead focus on survival times and leverage survival analysis techniques, specifically Cox proportional hazards models, to look for differences in survival times of enforcers versus non-enforcers.

@@

Unobserved Confounders: 
We further note several additional difference between the enforcers and the non-enforcer groups that were not considered in the @Popkin2023.  As an example, the number of average number of goals scores between the two groups significantly different for both definitions of enforcers.  The average number of goals scores for E-F enforcers  vs non-enforcers was `r summ$mngoals[summ$EF]` vs `r summ$mngoals[!summ$EF]` (p-value: `r ifelse(round(t.test(dat$G~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EF)$p.value,2))`).  For E-P enforcers, the average number of goals for enforcers was `r summEP$mngoals[summEP$EP]` vs `r summEP$mngoals[!summEP$EP]` (p-value: `r ifelse(round(t.test(dat$G~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EP)$p.value,2))`).

```{r echo = FALSE}
#Enforcer by country
epcountry <- dat %>% mutate(northamerica = ifelse(Ctry %in% c("USA","CAN"),"North America","Other")) %>% group_by(EP, northamerica) %>% summarize(n = n()) %>% pivot_wider(names_from = northamerica, values_from = n) %>% mutate(p = `North America`/(`North America`+ Other))

#prop.test(epcountry$`North America`,epcountry$Other+epcountry$`North America`)

#Enforcer by country
efcountry <- dat %>% mutate(northamerica = ifelse(Ctry %in% c("USA","CAN"),"North America","Other")) %>% group_by(EF, northamerica) %>% summarize(n = n()) %>% pivot_wider(names_from = northamerica, values_from = n)  %>% mutate(p = `North America`/(`North America`+ Other))

#prop.test(efcountry$`North America`,efcountry$Other+efcountry$`North America`)

#How many players are in each group
#table(dat$EP, dat$EF)

Perhaps more interestingly, we find highly significant differences in geographic region of enforcers vs non-enforcers for both definitions.  In the enforcer group `r paste0(100*round(efcountry$p[efcountry$EF],4),"\\%")` and `r paste0(100*round(epcountry$p[epcountry$EP],4),"\\%")` of those groups are from North America (i.e. United States or Canada) for E-F and E-P, respectively.  For non-enforcers, North American players make up only `r paste0(100*round(efcountry$p[!efcountry$EF],4),"\\%")` for E-F and `r paste0(100*round(epcountry$p[!epcountry$EP],4),"\\%")` for E-P.  For both definitions of enforcer, these differences are significant with p-values for both definitions less than 0.0001.

While we do not address this imbalance here, as we focus on closely following the work of @Popkin23, it is worth noting the substantial imbalance in geographic origin (i.e. North American vs outside of North America) of enforcers versus non-enforcer.
